---
title: "605_Week15_Assignment"
author: "Jeyaraman Ramalingam"
date: "5/10/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Taylor Series

The formula for Taylor Series is below

$$f(x)= \sum_{n=0}^{\infty} \dfrac{f^{n} (a)}{n!}(x-a)^n \\$$
$$ = f(a)+f`(a)(x-a)+\dfrac{f``(a)}{2!}(x-a)^2+\dfrac{f```(a)}{3!}(x-a)^3+\dfrac{f````(a)}{4!}(x-a)^4 \\$$

We can appy the formula for all the function mentioned 

## f(x) = 1/(1-x)

First step is to calculate the coefficients at a=0

$$f(x) = \dfrac{1}{(1-x)} = (1-x)^{-1}$$
$$ f(0) = (1-x)^{-1}$$
$$ f`(0) = (1-x)^{-2}=1$$
$$ f``(0) = 2(1-x)^{-3}=2$$
$$ f```(0) = 6(1-x)^{-4}=6$$
$$ f````(0) = 24(1-x)^{-4}=24$$
$$ f^{n}(x) = n!(1-x)^{-n}$$




Substituting the above in taylor series equation we get

$$Tn(x)=1+x+\dfrac{2}{2!} x^2+\dfrac{6}{3!} x^3+\dfrac{24}{4!} x^4+...$$



$$f(x)=1+x+x^2+x^3+x^4$$
Verify the coeffiecients using R function

```{r taylor_1,eval=TRUE}
library(pracma)

f <- function(x) 1/(1-x)

p <- round(taylor(f, 0, 4),2)
p

```

The values are matching with the coefficients in taylor series equation

Plot the Taylor Series Function 

```{r plot_taylor_1,eval=TRUE}

library(pracma)

f <- function(x) 1/(1-x)

p <- round(taylor(f, 0, 4),2)
p

s<- seq(-1.0, 1.0,0.1)
f_s <- f(s)

plt_taylor <- polyval(p, s)
plot(s, plt_taylor,type="l",col="red")
```

## f(x) = exp(x)


First step is to calculate the coefficients at a=0

$$f(x) = e^x $$
$$ f`(0) = e^1=1$$
$$ f``(0) = e^2=1$$
$$ f```(0) = e^2=1$$
$$ f````(0) = e^2=1$$
$$ f^{n}(x) = e^{n}$$




Substituting the above in taylor series equation we get

$$Tn(x)=1+x+\dfrac{x^2}{2!} +\dfrac{x^3}{3!} +\dfrac{x^4}{4!} +...$$

Verify the coeffiecients using R function

```{r taylor_2,eval=TRUE}
library(pracma)

f <- function(x) exp(x)

p <- round(taylor(f, 0, 4),2)
p

```

Plot the Taylor Series Function 

```{r plot_taylor_2,eval=TRUE}
library(pracma)

f <- function(x) exp(x)

p <- round(taylor(f, 0, 4),2)
p
s<- seq(-1.0, 1.0,0.1)
f_s <- f(s)

plt_taylor <- polyval(p, s)
plot(s, plt_taylor,type="l",col="red")
```

## f(x) = ln(1+x)

First step is to calculate the coefficients at a=0

$$f(x) = ln(1+x), f`(x)=\dfrac{1}{(1+x)}=(1+x)^{-1} $$
$$ f(0) = ln(1)=0$$
$$ f`(0) = (1+0)^-1=1$$
$$ f``(0) = -1(1+0)^-2=-1$$
$$ f```(0) = 2(1+0)^-3=2$$
$$ f````(0) = -6(1+0)^-4=-6$$
$$ f^{n}(x) = (-1)^{n+1}(n-1)!$$




Substituting the above in taylor series equation we get

$$Tn(x)=0+x-\dfrac{x^2}{2!} +2\dfrac{x^3}{3!} -6\dfrac{x^4}{4!} +...$$
$$Tn(x)=0+x-\dfrac{1}{2}x^2 +\dfrac{1}{3}x^3-\dfrac{1}{4}x^4$$


```{r taylor_3,eval=TRUE}
library(pracma)

f <- function(x) log(1+x)

p <- round(taylor(f, 0, 4),2)
p

```
Plot the Taylor Series Function 

```{r plot_taylor_3,eval=TRUE}
library(pracma)

f <- function(x) log(1+x)

p <- round(taylor(f, 0, 4),2)
p


s<- seq(-1.0, 1.0,0.1)
f_s <- f(s)

plt_taylor <- polyval(p, s)
plot(s, plt_taylor,type="l",col="red")
```

The values matches with the coeffieicients obtained in taylor series