---
title: "605_Week16_JR_Assignment"
author: "Jeyaraman Ramalingam"
date: "5/17/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r reg_line,eval=TRUE}
x <- c(5.6,6.3,7,7.7,8.4)
y <- c(8.8,12.4,14.8,18.2,20.8)

res<-lm(y~x)
res
summary(res)
plot(x,y)
abline(res,col="blue")
```

The Output from the linear regression model shows that the regression line can be written as below

$$-14.8+4.257x$$

## Problem 2
Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form
( x, y, z ). Separate multiple points with a comma.

$$f(x,y)=24x-6xy^{2}-8y^{3}$$

To Solve this , we need to write derivatives and then substitute the values to get the saddle points.


$$f(x,y)=24x-6xy^{2}-8y^{3}$$
$$fx=24-6y^{2}$$
$$fy=-12xy-24y^{2}$$
$$fxx=0$$
$$fyy=-12x-48y$$
$$fxy=-12y$$
$$fyx=-12y$$


Lets start with fx=0

$$24-6y^{2}=0$$
$$y=2,y=-2$$

Substituting the y-values in fy equation we get x = 4, x=-4

Substituting both x-values and y-values in the original equation we get the values 64, -64

Hence the Saddle Points are

(4,-2,64) (-4,2,-64)

## Problem 3

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she
sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81  21x + 17y units of the
"house" brand and 40 + 11x  23y units of the "name" brand.


###Step 1. Find the revenue function R ( x, y ).

$$R(x,y)=x(81-21x+17y)+y(40+11x-23y)$$
$$R(x,y)=81x-21x^{2}+17xy+40y+11xy-23y^{2}$$
$$R(x,y)=81x-21x^{2}+28xy+40y-23y^{2}$$
Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

The Value of x is House Brand which is 2.3 $ and the value for name brand is 4.10$

x = 2.3 , y =4.1

Substiuting the values in revenue function would provide us 

```{r solve,eval=TRUE}
rev <- function(x,y) {81*x-21*x^2+28*x*y+40*y-23*y^2}
rev
x=2.3
y=4.1
rev(x,y)
```

The Output is 116.62 and the total Revenue is ~ 117 dollars.

## Problem 4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by 
$$C(x,y)=\dfrac {1}{6}x^{2}+\dfrac {1}{6}y^{2} +7x+25y+700$$
where x is the number of
units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in
each plant to minimize the total weekly cost?


To Solve this problem we have x+y=96 , where x being the plant in LA and y being the plant in Denver

y=x-96

Substiuting the value of y in equation we will get 

$$\dfrac {1}{6}x^{2}+\dfrac {1}{6}(96-x)^{2} +7x+25(96-x)+700$$
After calculating derivatives we will get 

$$f`(x)=\dfrac {2}{3}x-50$$

This shows that the value of x is 75 and substiuting the x value in y=96-x=21 , the y-value is 21.

This means that the production units to minimize the total cost in LA is 75 and in Denver is 21.

## Problem 5

$$\underset{R}{\iint} (e^{8x+3y})dA; R:2\leq x\leq 4 , 2\leq y\leq 4$$

$$=\int_{2}^{4} (e^{8x})dx\int_{2}^{4} (e^{3y})dy$$

Solving the above integrals separately we get the below equation

$$=\dfrac {1}{24}(e^{32}-e^{16})(e^{12}-e^{6})$$

