---
title: "605_Week12_Assignment_JR"
author: "Jeyaraman Ramalingam"
date: "4/19/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction

The dataset is from R preloaded libraries and it consists of speed / distance values.

## Load datasets
```{r df,eval=TRUE}
library(ggplot2)
df<-get("cars")
df
```
## Correlation Coefficient 

Correlation coefficient provides us the relationship between the variables. If it is greater then a strong relationship exists and if it is low then the relationship is weak or none.

```{r correlation,eval=TRUE}
cor_df<-cor(df$speed, df$dist)
cor_df  
```

The Correlation coefficient is ~ 0.8 and this shows speed has strong impact over stopping distance. 

## Linear Model

```{r lm_bmi}


lm_speed <- lm(dist ~ speed, data = df)

lm_speed

summary(lm_speed)


ggplot(data = df, aes(x = speed, y = dist)) + 
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = FALSE)

```

## Plot Residuals

It is necessary to plot residuals and see if there are any variances.

```{r hist_bmi,eval=TRUE}
hist(lm_speed$residuals)
```

```{r lm_plot,eval=TRUE}
qqnorm(lm_speed$residuals)
qqline(lm_speed$residuals)  
plot(lm_speed$residuals ~ df$dist)

par(mfrow=c(2,2))
plot(lm_speed)
```

## Conclusion

We can conclude by saying that the data satisfies the conditions of linear model. Below are explanations.

**Linearity** - The data is linear as the residuals are following theoretical line.

**Homoscedasticity** - From Scale location plot we can see that the line is horizontal and not angled. This explains the homoscedasticity is satisfied.

**Independence** - Speed and Stopping distance are both independent variables

**Normality** - By looking at the QQ Plot  we can see that the data is distributed normally

