---
title: "605_JR_Week14_Homework"
author: "Jeyaraman Ramalingam"
date: "5/3/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Use integration by substitution to solve the integral below

$$ \int_{}^{}4e^{-7x}dx$$

Let u = -7x then 

du/dx=-7 

Integration equation will become

$$ \int_{}^{}4e^{u}\dfrac{du}{-7}$$
$$ = \dfrac{-4}{7}*\int_{}^{}4e^{u}du$$
$$ = \dfrac{-4}{7}*e^u + c$$
Substitute u = -7x

$$ = \dfrac{-4}{7}*e^-7x + c$$

The Answer is $$\dfrac{-4e^-7x}{7} + c$$

## Question 2

Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of

$$\dfrac{dN}{dt}=-\dfrac{3150}{t^4}-220$$ 

bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a
function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

Lets start with simplyfying the equation

$$\dfrac{dN}{dt}=-3150t^-4-220$$ 
Integration of above equation will return the function N(t)

$$N(t)=\int_{}^{}-3150t^-4-220 dt$$
$$N(t)=\dfrac{-3150t^-3}{-3}-220t +c$$
$$N(t)={1050t^-3}-220t +c$$
To find the constant c of the equation we need to use the equation

From the question N(t)=6530, t=1


$$6530 = 1050(1)^-3-220(1)+c$$
$$6530 = 1050-220+c$$
$$c = 6530 - 830$$
$$c = 5700$$

Apply c in the N(t) function to get N(t) for the given problem

$$N(t)={1050t^-3}-220t +5700$$

## Question 3

Find the total area of the red rectangles in the figure below, where the equation of the line is $$f(x)=2x-9$$

The red rectangles lower limit is 4.5
The red rectangles upper limit is 8.5

Integration of the function with the above limits would provide the answer.

```{r area,eval=TRUE}
gn_funct <- function(x) {2*x-9}
area_fn <- integrate(gn_funct,4.5,8.5)
area_fn
```

The answer is 16 with absolute error < 1.8e-13

## Question 4

Find the area of the region bounded by the graphs of the given equations.

$$y=x^2-2x-2,y=x+2$$

Enter your answer below.


Given two equations 

$$y = x^2-2x-2, y = x + 2$$


Solving the above equation will provide us the intercept points of the function

x = -2, x = 5

By using the above two points we can plot the graph in R

```{r plot,eval=TRUE}
limits <- seq(-2,5,0.5)

y1 <- function(x) {x^2-2*x-2}
y2 <- function(x) {x+2}
plot(limits, y1(limits), type='l', col="green", xlab="x-Axis", ylab="y-Axis")
lines(limits, y2(limits), type='l', col="red", xlab="x-Axis", ylab="y-Axis")
abline(h=0)

```

To find the areas we need to find the intersecting points

The points are -1, 4

Integrating the function with the above limits would provide the areas

```{r area_plot_1,eval=TRUE}
gn_funct <- function(x) {x^2-2*x-2}
area_fn <- integrate(gn_funct,-1,4)
area_fn
```

The area for the plot $$y=x^2-2x-2$$ is 

-3.333333 with absolute error < 1.2e-13

```{r area_plot_2,eval=TRUE}
gn_funct <- function(x) {x+2}
area_fn <- integrate(gn_funct,-1,4)
area_fn
```


The area for the plot $$y=x+2$$ is 

17.5 with absolute error < 1.9e-13

Total Area = 17.5+3.3 = ~20


## Question 5

A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year.
There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize
inventory costs.


$$C = 8.25*x+3.75*\dfrac {\dfrac {110}{x}}{2}$$

Simplyfying above would give us

$$C = 8.25*x+\dfrac{206.25}{x}$$

$$C' = 8.25-\dfrac{206.25}{x^2}$$

Let C'=0

$$8.25-\dfrac{206.25}{x^2}=0$$
$$8.25x^2-206.25=0$$
$$x^2=206.25/8.25$$
$$x^2=25$$
$$x=5$$

Number of orders to minimize inventory costs is 5 and lot size is 110/5 = 22.

## Question 6

Use integration by parts to solve the integral below.

$$ \int_{}^{}ln(9x)x^6dx$$


The Integration by parts formula is 

$$ \int_{}^{}udv$$ = uv - $$ \int_{}^{}vdu$$

$$ u = ln(9x) $$
$$ dv = x^6dx $$
$$ du/dx = 1/x $$
$$ v = \int_{}^{}x^6dx $$


Substituting above in the formula would give us

$$ \int_{}^{}ln(9x)x^6dx = ln(9x)\int_{}^{}x^6dx -  \int_{}^{}\int_{}^{}x^6dxdx/x$$

Solving the above equation would provide us the final answer

$$\dfrac{x^7}{7}(ln(9x)-\dfrac{1}{7})$$

## Question 7

Determine whether f ( x ) is a probability density function on the interval [1, e^6] . If not, determine the value of the
definite integral.

$$ f(x)= dfrac{1}{6x}$$

Lets test this by integrating the function with the limits

```{r integ_test,eval=TRUE}
gn_funct <- function(x) {1/(6*x)}
area_fn <- integrate(gn_funct,1,exp(6))
area_fn
```

The value is 1 which shows the function is PDF within the limits.

