---
title: "605_Week13_JR_Assignment"
author: "Jeyaraman Ramalingam"
date: "4/26/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

## Introduction

The attached who.csv dataset contains real-world data from 2008. The variables included follow.

Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
TotExp: sum of personal and government expenditures.

```{r df,eval=TRUE}
library(RCurl)
library(ggplot2)

file_url <- getURL("https://raw.githubusercontent.com/jey1987/DATA605/master/who.csv")
df_input <- read.csv(text=file_url,header=TRUE)
```

## Question 1

Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss
whether the assumptions of simple linear regression met.

### Plots


```{r qn_1_scatter,eval=TRUE}
plot(df_input$LifeExp~df_input$TotExp)
```

```{r qn_1_lm,eval=TRUE}

lm_life <- lm(df_input$LifeExp~df_input$TotExp)
summary(lm_life)

par(mfrow=c(2,2))
plot(lm_life)
abline(lm_life)
```

### Measure

**F-Statistics = 65.26**

**R-Square = 0.2537**

**Standard Error = 9.371**

**p-values = 7.714e-14**

### Interpretation

Based on the following conditions of Linear Regression, We can conclude that the data is not providing strong relationship.


Linearity - The data is non-linear as the residuals are not following theoretical line.

Homoscedasticity - From Scale location plot we can see that the line is not horizontal and angled. This explains the homoscedasticity is not satisfied.

Independence - Life Expectancy and Total Expenditure are both independent variables

Normality - By looking at the QQ Plot we can see that the data is not distributed normally

## Question 2

Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06
power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r
re-run the simple regression model using the transformed variables. Provide and interpret the F
statistics, R^2, standard error, and p-values. Which model is "better?"

### Plots 

```{r qn_2_scatter,eval=TRUE}
df_input$LifeExp = df_input$LifeExp^4.6
df_input$TotExp = df_input$TotExp^0.06

plot(df_input$LifeExp~df_input$TotExp)
```

```{r qn_2_plot,eval=TRUE}

lm_life <- lm(df_input$LifeExp~df_input$TotExp)
summary(lm_life)

par(mfrow=c(2,2))
plot(lm_life)
abline(lm_life)
```

### Measure

**F-Statistics = 507.7**

**R-Square = 0.7283**

**Standard Error = 90490000**

**p-values = < 2.2e-16**

### Interpretation

Based on the following conditions of Linear Regression, We can conclude that the data is not providing strong relationship.


Linearity - The data is linear as the residuals are following theoretical line.

Homoscedasticity - From Scale location plot we can see that the line is horizontal and not angled. This explains the homoscedasticity is satisfied.

Independence - Life Expectancy and Total Expenditure are independent variables

Normality - By looking at the QQ Plot we can see that the data is distributed normally

By comparing the conditions of linear regression, We can say that Model 2 is better than 1


## Question 3

Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life
expectancy when TotExp^.06=2.5. 

```{r qn_3_predict,eval=TRUE}
mod <- lm(LifeExp~TotExp,data=df_input)
newd <- data.frame(TotExp=c(1.5,2.5))
predict(mod, newd, interval="predict")^(1/4.6)
```

The Life Expectancy is forecasted at 63 years.

## Question 4

Build the following multiple regression model and interpret the F Statistics, R^2, standard error,
and p-values. How good is the model?

LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

```{r qn_4_mod,eval=TRUE}

file_url <- getURL("https://raw.githubusercontent.com/jey1987/DATA605/master/who.csv")
df_input <- read.csv(text=file_url,header=TRUE)

mod_new <- lm(LifeExp ~ PropMD + TotExp + PropMD * TotExp,data=df_input)
summary(mod_new)

par(mfrow=c(2,2))
plot(mod_new)
abline(mod_new)
```

### Measure

**F-Statistics = 34.49**

**R-Square = 0.3471**

**Standard Error = 8.765**

**p-values = < 2.2e-16**

### Interpretation

Based on the following conditions of Linear Regression, We can conclude that the data is not providing strong relationship.


Linearity - The data is non-linear as the residuals are not following theoretical line.

Homoscedasticity - From Scale location plot we can see that the line is not horizontal and angled. This explains the homoscedasticity is not satisfied.

Independence - Life Expectancy and Total Expenditure are both independent variables

Normality - By looking at the QQ Plot we can see that the data is not distributed normally


## Question 5

Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why
or why not?

```{r qn_5,eval=TRUE}
newd <- data.frame(PropMD=0.03,TotExp=14)
predict(mod_new, newd, interval="predict")
```

The Life Expectancy is forecasted at 107 years, this seems not possible considering the average life of humans is ~ 70.